 Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021.

### 2023-02-03 outputs/paper/Paper 1.qmd
Toronto Victimization: Seventh commit, including all updated inputs, outputs and scripts

---
title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021."
author: Sofia Sellitto
date: February 3, 2023
date-format: long
format:
  pdf: 
    documentclass: article 
    geometry: margin = 1in
    abstract: In the city of Toronto, citizens have become accustomed to fearing for their safety as crime appears to have increased in recent years. In this paper, I analyze the number of victims of crimes against the person in the City of Toronto from 2014 - 2021 to investigate if the number of Torontonians victimized by crime has increased over time and if sex differences are apparent across various categories of crime. By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time. Sexual violence appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female. These results have implications for how Torontonians view their safety and consider how sex impacts their predisposition to certain crimes. 
toc: true
toc-depth: 4
theme: cosmo
bibliography: bibliography.bib
---

@@

In this paper, I utilize Open Data Toronto [@opendatatoronto] to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. This is to investigate how the number of victims of crime in Toronto has changed over time, if it is increasing, which categories of crimes are most prevalent and if sex differences are apparent across different forms of crime. I will commence with a discussion of my data source and its features. I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violence" and "other", across 2014-2019. Next, I will move towards an analysis of how the number of victims has changed according to the category of crime, over time. Following, I will conduct an investigation of how the number of victimization reports for each category of crime, differs according to the sex of the victim. The paper will culminate with a discussion of my findings, limitations and future research possibilities.

@@

\
Included within the crime category subtype are (1) assault, (2) robbery, (3) sexual violation and (4) other. To further discuss these categorizations, in Canada, assault is classified as intentionally applying force to an individual indirectly or directly without consent. This can then include assault to cause bodily harm, with a deadly weapon, aggravated assault or attempted murder. In Canada, robbery is defined as stealing for extortion of the article stolen and can include the use of violent threats, personal violence or with a weapon. In Ontario, sexual violence describes any physical or psychological violence conducted or targeted in a sexual means. Unfortunately, no information was provided by the City of Toronto nor the Toronto Police Service Annual Statistical Report (ASR) on the nature of the "other" category. We are forced to assume that this may include break and enter, homicide, domestic violence or various forms of crimes against the person.

@@

See **Figure 1**

@@

"Sexual Violence" had the least number of victims of crime from 2014 - 2021. However, contrary to the others, sexual violence saw its peak in 2021 with 2568 victims. From 2014 onward, the number of victims gradually increased until it peaked in 2021.

In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "Sexual Violence". Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively. Using the data set from Table 3, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.

\
See **Figure 2**

@@

In drawing your attention to Figure 2 contained in the "Assault", on the top left, it is apparent that from 2014 - 2021, more male Torontonians were victims of assault than women. In its peak year, 2017, of the 18819 victims, 8679 were women and 10140 were men. This would mean that in 2017, there was 1461 more male than female victims, with a difference of 15.53%.

@@

\
Lastly, "Sexual Violence", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violence than men. This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021, being female, and 334 being male. This means a difference of 147.9%.

@@

In this paper, I have utilized The City of Torontos open data portal [@opendatatoronto], supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed, if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent. By utilizing the Coding language R [@R] I was able to import, clean and asses the raw data made available by the City. In doing so I discovered that for the sub-category of crimes, assault, robbery, and other, the number of Torontonian victims peaked in the years 2019, 2017, and 2015 respectively and decreased until the end of the data collection date (2021). The most prevalent crime that Torontonians were victims of was assault, maintaining the highest cumulative average across 2014-2021. Contrarily, sexual violence saw its peak in 2021, meaning it has not decreased over time. In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violence had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%. Robbery maintains the second largest difference between male and female victims, skewing towards men with a difference of 101.35%. "Other", had the third largest difference, with more female victims than male. Assault had the least large sex difference, with slightly more male victims than feel with a difference of 15.53%.

@@

Because of the nature of the data set, my analysis is limited in many regards. To begin with the source, Torontos Police Service Annual Statistical Report (ASR), it is possible that data was misrepresented, incorrectly input, missing, duplicated or potentially falsified. While it is important to avoid speculation about these statistics' validity, I believe it is necessary to highlight possibilities for error. Another vital point to acknowledge is the number of individuals who may have fallen victim to a crime against the person and failed to report it or those who intended to file a report but were dismissed. The latter is an extremely apparent phenomenon for women who attempt to report sexual violence. While this data represents the number of individuals who have fallen victim to crimes against the person across 2014 - 2021, it cannot account for unfilled and dismissed reports. Meaning that the number of victims could potentially be far higher than portrayed in the analysis. Regarding the data set itself, removing missing and unknown data will impact the analysis in various regards. More specifically, concerning the variable sex including the categories "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination. Removing this data could have prevented insightful analysis of non-binary and Agender individuals.

>>>>

---
title: " Sin and The City: Investigating the Number and Sex of Victims of Crimes Against the Person in Toronto Across 2014-2021."
author: Sofia Sellitto
date: February 3, 2023
date-format: long
format:
  pdf: 
    documentclass: article 
    geometry: margin = 1in
    abstract: In the city of Toronto, citizens have become accustomed to fearing for their safety as crime appears to have increased in recent years. In this paper, I analyze the number of victims of crimes against the person in the City of Toronto from 2014 - 2021 to investigate if the number of Torontonians victimized by crime has increased over time and if sex differences are apparent across various categories of crime. By examining the data I found that in all categories of crime (assault, robbery and other) but sexual assault, the number of annual victims has decreased over time. Sexual violation appears to have increased in 2021 maintaining the largest sex difference, skewing drastically female. These results have implications for how Torontonians view their safety and consider how sex impacts their predisposition to certain crimes. 
toc: true
toc-depth: 4
theme: cosmo
bibliography: bibliography.bib
---

@@

In this paper, I utilize Open Data Toronto [@opendatatoronto] to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. This is to investigate how the number of victims of crime in Toronto has changed over time, if it is increasing, which categories of crimes are most prevalent and if sex differences are apparent across different forms of crime. I will commence with a discussion of my data source and its features. I will then conduct a breakdown of the average number of victimization reports for the sub-categories "assault", "robbery", "sexual violation" and "other", across 2014-2019. Next, I will move towards an analysis of how the number of victims has changed according to the category of crime, over time. Following, I will conduct an investigation of how the number of victimization reports for each category of crime, differs according to the sex of the victim. The paper will culminate with a discussion of my findings, limitations and future research possibilities.

@@

\
Included within the crime category subtype are (1) assault, (2) robbery, (3) sexual violation and (4) other. To further discuss these categorizations, in Canada, assault is classified as intentionally applying force to an individual indirectly or directly without consent. This can then include assault to cause bodily harm, with a deadly weapon, aggravated assault or attempted murder. In Canada, robbery is defined as stealing for extortion of the article stolen and can include the use of violent threats, personal violence or with a weapon. In Ontario, sexual violation describes any physical or psychological violence conducted or targeted in a sexual means. Unfortunately, no information was provided by the City of Toronto nor the Toronto Police Service Annual Statistical Report (ASR) on the nature of the "other" category. We are forced to assume that this may include break and enter, homicide, domestic violence or various forms of crimes against the person.

@@

@@

"sexual violation" had the least number of victims of crime from 2014 - 2021. However, contrary to the others, sexual violation saw its peak in 2021 with 2568 victims. From 2014 onward, the number of victims gradually increased until it peaked in 2021.

In representing how sex impacts the victims of crimes, I chose to display this data using 4 bar charts based on the categories "Assault", "Other", "Robbery" and "sexual violation". Each plot assesses the number of victims across 2014 - 2021 and the number of victims that were male (M) and female (F), respectively. Using the data set from Table 3, which was previewed above, I determined the number of victims that were male and female for the peak years of each sub-category of crime and calculated percentage differences.

@@

```
In drawing your attention to Figure 2 contained in the "Assault", on the top left, it is apparent that from 2014 - 2021, more male Torontonians were victims of assault than women. In its peak year, 2017, of the 18819 victims, 8679 were women and 10140 were men. This would mean that in 2017, there was 1461 more male than female victims, with a difference of 15.53%.

@@

Lastly, "sexual violation", on the bottom right, shows the most apparent sex difference from 2014-2021, where more female Torontonians were victims of sexual violation than men. This sex difference was the most defined across all the sub-categories, with 2244 of the total 2568 victims in its peak year, 2021, being female, and 334 being male. This means a difference of 147.9%.

@@

In this paper, I have utilized The City of Torontos open data portal [@opendatatoronto], supplied by the Toronto Police Service Annual Statistical Report (ASR) to assess the number of victims of crimes against the person in Toronto across 2014 - 2021. The primary objective of this paper was to assess if the number of Torontonians victimized by crime has changed, if it has increased over time, which categories of crimes against the person are most prevalent and if sex differences are apparent. By utilizing the Coding language R [@R] I was able to import, clean and asses the raw data made available by the City. In doing so I discovered that for the sub-category of crimes, assault, robbery, and other, the number of Torontonian victims peaked in the years 2019, 2017, and 2015 respectively and decreased until the end of the data collection date (2021). The most prevalent crime that Torontonians were victims of was assault, maintaining the highest cumulative average across 2014-2021. Contrarily, sexual violation saw its peak in 2021, meaning it has not decreased over time. In assessing how sex affects the number of victims for each sub-category of crime, it was apparent that sexual violation had the largest difference between male and female victims, skewing drastically female, maintaining a difference of 147.97%. Robbery maintains the second largest difference between male and female victims, skewing towards men with a difference of 101.35%. "Other", had the third largest difference, with more female victims than male. Assault had the least large sex difference, with slightly more male victims than feel with a difference of 15.53%.

@@

Because of the nature of the data set, my analysis is limited in many regards. To begin with the source, Torontos Police Service Annual Statistical Report (ASR), it is possible that data was misrepresented, incorrectly input, missing, duplicated or potentially falsified. While it is important to avoid speculation about these statistics' validity, I believe it is necessary to highlight possibilities for error. Another vital point to acknowledge is the number of individuals who may have fallen victim to a crime against the person and failed to report it or those who intended to file a report but were dismissed. The latter is an extremely apparent phenomenon for women who attempt to report sexual violation. While this data represents the number of individuals who have fallen victim to crimes against the person across 2014 - 2021, it cannot account for unfilled and dismissed reports. Meaning that the number of victims could potentially be far higher than portrayed in the analysis. Regarding the data set itself, removing missing and unknown data will impact the analysis in various regards. More specifically, concerning the variable sex including the categories "male", "female" and "unknown", I was forced to remove "unknown" as further insights were not provided on this denomination. Removing this data could have prevented insightful analysis of non-binary and Agender individuals.
